---
title: "Custom sh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary 

A custom function for calling `gcloud` SDK from Rmarkdown via OS command `system2()` 

## Create function

Create function for calling `gcloud` SDK from Rmarkdow:

```{r function, warning=FALSE}
library(withr)
library(glue)

sh <- function(cmd, args = c(), intern = FALSE) {
  with_path(path.expand("~/google-cloud-sdk/bin/"), {
    if (is.null(args)) {
      cmd <- glue(cmd)
      s <- strsplit(cmd, " ")[[1]]
      cmd <- s[1]
      args <- s[2:length(s)]
    }
    ret <- system2(cmd, args, stdout = TRUE, stderr = TRUE)
    if ("errmsg" %in% attributes(attributes(ret))$names) cat(attr(ret, "errmsg"), "\n")
    if (intern) return(ret) else cat(paste(ret, collapse = "\n"))
  })}
```

## Authenticate with Google Cloud and Vertex AI 

Authenticate with Google Cloud and Vertex AI via `googleAuthR` package and `googleCloudVertexAIR`: 

```{r auth, warning=FALSE}
library(googleAuthR)
library(googleCloudVertexAIR)

options(googleAuthR.scopes.selected = "https://www.googleapis.com/auth/cloud-platform")

gar_auth_service(json_file = Sys.getenv("GAR_SERVICE_JSON"))
```

Set `gcloud` project: 

```{r}
sh("gcloud config set project {gcva_project_get()}")
```

List `gcloud` configuration to verify: 

```{r results='hide'}
sh("gcloud config list")
```


## Acknowledgements 

Special thanks to [Fabian Hirschmann](https://github.com/fhirschmann, the original author of this source code. 
