---
title: "Training and deploying a custom model"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{customtabularonline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

Emulation of

* blog post: [Train and deploy ML models with R and plumber on Vertex AI | Google Cloud Blog](https://cloud.google.com/blog/products/ai-machine-learning/train-and-deploy-ml-models-with-r-and-plumber-on-vertex-ai)
* notebook: [vertex-ai-samples/get\_started\_vertex\_training\_r\_using\_r\_kernel.ipynb at main · GoogleCloudPlatform/vertex-ai-samples](https://github.com/GoogleCloudPlatform/vertex-ai-samples/blob/main/notebooks/community/ml_ops/stage2/get_started_vertex_training_r_using_r_kernel.ipynb)

[Create](https://cloud.google.com/vertex-ai/docs/training/create-training-pipeline) a `TrainingPipeline`  that runs a `CustomJob` and imports the resulting artifacts as a `Model`.

## Setup

### Installation

Run the following chunk to install `googleCloudVertexAIR` and the other required R packages to complete this tutorial (checking to see if they are installed first and only install if not already):

```{r install_packages, eval=FALSE}
required_packages <- c("remotes", "googleAuthR", "withr")
missing_packages <- required_packages[!(required_packages %in%
                                          installed.packages()[,"Package"])]
if(length(missing_packages)) install.packages(missing_packages)
# remotes::install_github("justinjm/googleCloudVertexAIR") # run first time
# options(googleAuthR.verbose = 0) # set when debugging
```

```{r warning=FALSE}
library(withr) # for gcloud sdk: https://stackoverflow.com/a/64381848/1812363 
library(glue) 

sh <- function(cmd, args = c(), intern = FALSE) {
  with_path(path.expand("~/google-cloud-sdk/bin/"), {
    if (is.null(args)) {
      cmd <- glue(cmd)
      s <- strsplit(cmd, " ")[[1]]
      cmd <- s[1]
      args <- s[2:length(s)]
    }
    ret <- system2(cmd, args, stdout = TRUE, stderr = TRUE)
    if ("errmsg" %in% attributes(attributes(ret))$names) cat(attr(ret, "errmsg"), "\n")
    if (intern) return(ret) else cat(paste(ret, collapse = "\n"))
  }
  )
}
```

### .Renviron

Create a file called `.Renviron` in your project's working directory and use the following environemtn arguments:

* `GAR_SERVICE_JSON` - path to service account (JSON) keyfile downloaded before and copied
* `GCVA_DEFAULT_PROJECT_ID` - string of your GCP project you configured before
* `GCVA_DEFAULT_REGION` - region of GCP resorces that can be one of: `"us-central1"` or `"eu"`

e.g. your `.Renviron` should look like:

```
# .Renviron
GAR_SERVICE_JSON="/Users/me/auth/auth.json"
GCVA_DEFAULT_PROJECT_ID="my-project"
GCVA_DEFAULT_REGION="us-central1"
```

## Setup Google Cloud Project

TODO

## Authenticate your Google Cloud Account

```{r auth, warning=FALSE}
library(googleAuthR)
library(googleCloudVertexAIR)

options(googleAuthR.scopes.selected = "https://www.googleapis.com/auth/cloud-platform")

gar_auth_service(json_file = Sys.getenv("GAR_SERVICE_JSON"))
```

## Set global arguements

```{r gcva-global-arguements}
gcva_region_set(region = "us-central1")
gcva_project_set(projectId = Sys.getenv("GCVA_DEFAULT_PROJECT_ID"))

timestamp <- strftime(Sys.time(), "%Y%m%d%H%M%S")
paste0("timestamp: ", timestamp)

datasetDisplayName <- sprintf("california-housing-%s", timestamp)
paste0("datasetDisplayName: ", datasetDisplayName)

bucket_name <- paste0(gcva_project_get(), "-aip-", timestamp)
bucket_uri <- paste0("gs://", bucket_name)

paste0("bucket_name: ", bucket_name)
paste0("bucket_uri: ", bucket_uri)

private_repo <- "my-docker-repo"
image_name <- "vertex-r"  
image_tag <- "latest"
image_uri <- glue(
  "{gcva_region_get()}-docker.pkg.dev/{gcva_project_get()}/{private_repo}/{image_name}:{image_tag}"
)
image_uri 
```

## Create cloud storage bucket 

```{r}
sh("gcloud config set project {gcva_project_get()}")
```

```{r}
sh("gsutil mb -l {gcva_region_get()} -p {gcva_project_get()} {bucket_uri}")
```

## Create a private docker repository 

Your first step is to create your own Docker repository in Google Artifact Registry.

Run the gcloud artifacts repositories create command to create a new Docker repository with your region with the description "docker repository".

Run the gcloud artifacts repositories list command to verify that your repository was created.

```{r}
sh('gcloud artifacts repositories create {private_repo} --repository-format=docker --location={gcva_region_get()} --description="Docker repository"')
```


```{r}
sh("gcloud artifacts repositories list")
```

### Configure authentication to your private repo

Before you push or pull container images, configure Docker to use the `gcloud` command-line tool to authenticate requests to `Artifact Registry` for your region.


```{r}
sh("gcloud auth configure-docker {gcva_project_get()}-docker.pkg.dev --quiet")
```


## Create Dockerfile 

```{r}
dir.create("src", showWarnings=FALSE)
```


`.src/Dockerfile`

```{r}
Dockerfile = cat(
"FROM gcr.io/deeplearning-platform-release/r-cpu.4-1:latest

WORKDIR /root

COPY train.R /root/train.R
COPY serve.R /root/serve.R

# Install Fortran
RUN apt-get update
RUN apt-get install gfortran -yy

# Install R packages
RUN Rscript -e \"install.packages('plumber')\"
RUN Rscript -e \"install.packages('randomForest')\"

EXPOSE 8080
", file = "src/Dockerfile")
```

## Create training script 

`.src/train.R`

```{r}
cat(
'#!/usr/bin/env Rscript
library(tidyverse)
library(data.table)
library(randomForest)
Sys.getenv()

# The GCP Project ID
project_id <- Sys.getenv("CLOUD_ML_PROJECT_ID")

# The GCP Region
location <- Sys.getenv("CLOUD_ML_REGION")

# The Cloud Storage URI to upload the trained model artifact to
model_dir <- Sys.getenv("AIP_MODEL_DIR")

# Next, you create directories to download our training, validation, and test set into.
dir.create("training")
dir.create("validation")
dir.create("test")

# You download the Vertex AI managed data sets into the container environment locally.
system2("gsutil", c("cp", Sys.getenv("AIP_TRAINING_DATA_URI"), "training/"))
system2("gsutil", c("cp", Sys.getenv("AIP_VALIDATION_DATA_URI"), "validation/"))
system2("gsutil", c("cp", Sys.getenv("AIP_TEST_DATA_URI"), "test/"))

# For each data set, you may receive one or more CSV files that you will read into data frames.
training_df <- list.files("training", full.names = TRUE) %>% map_df(~fread(.))
validation_df <- list.files("validation", full.names = TRUE) %>% map_df(~fread(.))
test_df <- list.files("test", full.names = TRUE) %>% map_df(~fread(.))

print("Starting Model Training")
rf <- randomForest(median_house_value ~ ., data=training_df, ntree=100)
rf

saveRDS(rf, "rf.rds")
system2("gsutil", c("cp", "rf.rds", model_dir))
', file = "src/train.R")
```

## Create serving script 

`.src/serve.R`

```{r}
cat(
'#!/usr/bin/env Rscript
Sys.getenv()
library(plumber)

system2("gsutil", c("cp", "-r", Sys.getenv("AIP_STORAGE_URI"), "."))
system("du -a .")

rf <- readRDS("artifacts/rf.rds")
library(randomForest)

predict_route <- function(req, res) {
    print("Handling prediction request")
    df <- as.data.frame(req$body$instances)
    preds <- predict(rf, df)
    return(list(predictions=preds))
}

print("Staring Serving")

pr() %>%
    pr_get(Sys.getenv("AIP_HEALTH_ROUTE"), function() "OK") %>%
    pr_post(Sys.getenv("AIP_PREDICT_ROUTE"), predict_route) %>%
    pr_run(host = "0.0.0.0", port=as.integer(Sys.getenv("AIP_HTTP_PORT", 8080)))
', file = "src/serve.R")
```

## Build container and push to artifact registry 

```{r}
sh("gcloud services enable artifactregistry.googleapis.com")
```

```{r}
sh("cd src && gcloud builds submit --region={gcva_region_get()} --tag={image_uri} --timeout=1h")
```

## Create Dataset 

```{r}
data_uri <- "gs://cloud-samples-data/ai-platform-unified/datasets/tabular/california-housing-tabular-regression.csv"

dataset <- gcva_create_tabluar_dataset(
  displayName = datasetDisplayName,
  gcsSource = data_uri)
dataset
```

## Create custom container training job 

* Doc: [Create training pipelines > Custom Container > REST | Vertex AI | Google Cloud](https://cloud.google.com/vertex-ai/docs/training/create-training-pipeline#aiplatform_create_training_pipeline_custom_job_sample-drest)
* Python SDK: [Class CustomContainerTrainingJob | Python client library | Google Cloud](https://cloud.google.com/python/docs/reference/aiplatform/latest/google.cloud.aiplatform.CustomContainerTrainingJob)
* API: [REST Resource: projects.locations.trainingPipelines | Vertex AI | Google Cloud](https://cloud.google.com/vertex-ai/docs/reference/rest/v1/projects.locations.trainingPipelines#TrainingPipeline)


```{r}
job <- gcva_custom_container_training_job(
  displayName = image_name,
  containerUri = image_uri,
  command = c("Rscript", "train.R"),
  modelServingContainerCommand = c("Rscript", "serve.R"),
  modelServingContainerImageUri = image_uri
)
```

## Execute custom training job 

```r
model <- gcva_job_run(
  job = job,
  dataset = dataset,
  modelDisplayName = modelDisplayName,
  machineType = "n1-standard-4"
)

model$display_name #python sdk
model$resource_name  #python sdk
model$uri  #python sdk
```

## Create Endpoint 

[Overview of getting predictions on Vertex AI | Google Cloud](https://cloud.google.com/vertex-ai/docs/predictions/overview#model_deployment)

```r
endpoint <- gcva_create_endpoint(
    display_name = "California Housing Endpoint",
    project = PROJECT_ID,
    location = REGION
)
endpoint
```

## Deploy Model to Endpoint 

```r
gcva_deploy(
  model = model,
  endpoint = endpoint, 
  machine_type = "n1-standard-4"
)
```

## Test deployed model 


### Create sample input data 

```{r}
library(jsonlite)
df <- read.csv(text=sh("gsutil cat {data_uri}", intern = TRUE))
head(df, 5)

instances = list(instances=head(df[, names(df) != "median_house_value"], 5))
instances

json_instances = toJSON(instances)
json_instances
```

### submit request 

```r 
gcva_predict(
  instances = json_instances 
)
```

## Cleanup

TODO

### Undeploy 

```r
gcva_undeploy_all(endpoint = endpoint)
gcva_delete_endpoint()
```
### Deletes 

```r
gcva_delete_job()
```

```{r}
gcva_delete_dataset(displayName = datasetDisplayName)
```

local directory 

```{r}
unlink("src", recursive = TRUE) # will delete directory called 'mydir'
```

GCS bucket 

```{r}
delete_bucket <- TRUE 
if (delete_bucket == TRUE) sh("gsutil -m rm -r {bucket_uri}")
```
